### 2014-03-14 text/Chapter00.tex
Add dataflow example to introduction

\subsection{Context-awareness \#4: Access patterns in data-flow languages}

As an example, consider a language that allows us to get a value of a variable (representing
some changing data-source) \ident{x} versions back using the syntax $\ident{x}_{[a]}$. 
To track information about individual variables, we use a product-like operation $\times$ on tags 
to mirrors the product structure of variables. For example:
%
\begin{equation*}
?
\end{equation*}
%
The annotations $5 \times 10$ corresponds to the free-variable context $\ident{a}, \ident{b}$, denoting
that we need at most 5 and 10 past values of \ident{a} and \ident{b}, respectively. If we substitute 
\ident{c} for both \ident{a} and \ident{b}, we need to combine the context-requirements and take the
maximum of the requirements of the individual variables:
%
\begin{equation*}
\begin{array}{l}
?
\end{array}
\end{equation*}
%
This version of the calculus removes the non-determinism of the lambda abstraction from the previous
version. As we associate information with individual variables, lambda abstraction creates a function
that requires the context requirements associated with the variable that is being abstracted over.

For example, if we wrapped the earlier example above in a function taking \ident{a} (and using 
\ident{b} from the declaration-site) then the function would have context requirements $5$ -- that
is the number associated with the variable \ident{a}.

>>>>

\subsection{Context-awareness \#4: Checking array access patterns}

The last example leaves the topic of cross-platform and distributed computing. We focus on 
checking how arrays are accessed. This is a simpler version of the data-flow programming
examples used later in the thesis.

Consider a simple programming language with arrays where $n^{\textnormal{th}}$ element of an
array \ident{arr} is accessed using $\ident{arr}[n]$. Furthermore, we focus on performing local
transformations and we assume that the keyword $\kvd{cursor}$ returns the \emph{current} location
in the array.

The following example implements a simple one-dimensional cellular automata, reading from the
\ident{input} array and writing to \ident{output}:
%
\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{sum}=\ident{input}[\kvd{cursor} - 1] ~+~ \ident{input}[\kvd{cursor}] ~+~ \ident{input}[\kvd{cursor} + 1]\\[-0.25em]
\kvd{if}~\ident{sum}=2 ~||~ (\ident{sum}=1 ~\&\&~ \ident{input}[\kvd{cursor} - 1] = 0)\\[-0.25em] 
\kvd{then}~\ident{output}[\kvd{cursor}] \leftarrow 1 ~\kvd{else}~\ident{output}[\kvd{cursor}] \leftarrow 0

\end{array}
\end{equation*}
%
In this example, we use the term \emph{context} to refer to the values in the array around the 
current location provided by \kvd{cursor}. The interesting question is, how much of the context
(\ie~how far in the array) does the program access.

This is contextual information attached to individual (array) variables. 
In the above example, we want to track that \ident{input} is accessed in the range $\langle-1, 1 \rangle$
while \ident{output} is accessed in the range $\langle 0, 0 \rangle$. When calculating the ranges,
we need to be able to compose ranges $\langle-1,-1$, $\langle 0,0 \rangle$ and $\langle 1,1 \rangle$
(based on the accesses on the first line).

The information about access patterns can be used to efficiently compile the computation
(as we know which sub-range of the array might be accessed) and it also allows better handling
of boundaries. For example, wrap-around behaviour we could pad the \ident{input} with a known
number of elements from the other side of the array.
